### 2014-04-15 brains-protocol.tex
Rework presentation of measures

For a given feature $\mathfrak{F}$, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
Word substitution susceptibility (denoted $\mathfrak{S}_{\mathfrak{F}}(w)$) is computed as the ratio of the number of times $n_s(w)$ a word is substituted to the number of times $n_p(w)$ that word appears in a substitutable position.
We have:
$$\mathfrak{S}_{\mathfrak{F}}(w) \defeq \frac{n_s(w)}{n_p(w)}$$

Now averaging over all words having a given feature value $f$, we obtain the mean susceptibility for the feature value $f$:
\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
$$\left< \mathfrak{S}_{\mathfrak{F}} \right>_f = \left< \frac{n_s(w)}{n_p(w)} \right>_{\left\lbrace w | \mathfrak{F}(w) = f \right\rbrace}$$

@@

Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$.
Let us denote this quantity:
$$\Delta (\wstart,\warrival) \defeq \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart)$$

Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.
This quantity can be written:
$$\left< \Delta(\wstart,\warrival) \right>_f = \left< \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
The corresponding average quantity over all start words may be written:
\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word}, i.e. an arrival word chosen among semantically plausible though still random words.
\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}

>>>>

For a given feature $\phi$, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
Word substitution susceptibility is computed as the ratio of the number $s_w$ of times a word is substituted to the number $p_w$ of times that word appears in a substitutable position, i.e. $\sfrac{s_w}{p_w}$.

Now averaging over all words such that $\phi(w) = f$, we obtain the mean susceptibility for the feature value $f$:
\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
$$\sigma_{\phi}(f) \defeq \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$

@@

Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival) - \phi(\wstart)$.
Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
$$\Delta_{\phi}(f) \defeq \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \phi(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
The new quantity under $\mathcal{H}_0$ is:
\footnote{Note that $\phi(\warrival_{\mathcal{H}_0})$ is in fact a constant in this averaging, since by definition $\warrival_{\mathcal{H}_0}$ does not depend on $\wstart$.}
$$\Delta_{\phi, \mathcal{H}_0}(f) \defeq \left< \phi({\warrival_{\mathcal{H}_0}}) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \phi(\wstart) = f \right\rbrace}$$

We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word}, i.e. an arrival word chosen among semantically plausible though still random words.
\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}\rk{S: Then we must say the results are the same.}
